
To start we need to import the library. The library stores the comands wich are used to 
visualize the data.

```{r}
library(tidyverse)
```

here the Dataset is imported, so it is posible to acces the information from it. 

```{r}
info <- read.csv("qog_eureg_wide1_nov20.csv")
```

This Part is used to compare the developpment of school early leavers for the time scince 
2000 till 2019 for women and men.
To Start the library is loadet and the datadet is opend into the system. Then before 
visualisation the data gets filtert. To start that only the information of those rows is 
accesed wich havethe region name GERMANY. Secont the information is selected so that there 
is region name and the the year available. Further the limits are beeing set. Therefore 
we filter futhere that only the walues we need are available. They are also filtert fore 
the data from the "nuts0" part. then the seperat steps get combyned so there is only one 
variable is needed to acces the information.
The last part is the plot wich displayes the data. 

```{r comparison of erlyLevers for male and female}
library(tidyverse)

info <- read.csv("qog_eureg_wide1_nov20.csv")

germaninfo <- filter(info, region_name == "GERMANY")

specef <-  select(germaninfo, region_name, year)

filtered <- select(germaninfo, eu_eduleave_m_nuts0:eu_eduleave_f_nuts0)

filtered <- select(filtered, ends_with("nuts0"))

merged <- bind_cols(specef, filtered)

merged <- filter(merged, year >= 2000 & year < 2019)

ggplot(data = merged) +
  geom_line(mapping = aes(x = year, y = eu_eduleave_m_nuts0, color = "Female")) +
  geom_line(mapping = aes(x = year, y = eu_eduleave_f_nuts0, color = "Male")) +
  scale_color_manual(name = "Gender", values = c("Female" = "blue", "Male" = "red" ))
```



```{r regional educaton}
library(tidyverse)

info <- read.csv("qog_eureg_wide1_nov20.csv")

ger <- filter(info, region_name == "GERMANY")
fr <- filter(info, region_name == "FRANCE")
nl <- filter(info, region_name == "NETHERLANDS")
pl <- filter(info, region_name == "POLAND")

gerinfo <- select(ger, region_name, year)
frinfo <- select(fr, region_name, year)
nlinfo <- select(nl, region_name, year)
plinfo <- select(pl, region_name, year)

ger_fil <- select(ger, eu_empl_durtotal_nuts0:eu_neet_y1524t_nuts0)
fr_fil <- select(fr, eu_empl_durtotal_nuts0:eu_neet_y1524t_nuts0)
nl_fil <- select(nl, eu_empl_durtotal_nuts0:eu_neet_y1524t_nuts0)
pl_fil <- select(pl, eu_empl_durtotal_nuts0:eu_neet_y1524t_nuts0)

ger_fil <- select(ger_fil, ends_with("nuts0"))
fr_fil <- select(fr_fil, ends_with("nuts0"))
nl_fil <- select(nl_fil, ends_with("nuts0"))
pl_fil <- select(pl_fil, ends_with("nuts0"))

ger_merge <- bind_cols(gerinfo, ger_fil)
fr_merge <- bind_cols(frinfo, fr_fil)
nl_merge <- bind_cols(nlinfo, nl_fil)
pl_merge <- bind_cols(plinfo, pl_fil)

ger_merge <- filter(ger_merge, year >= 2000 & year < 2019)
fr_merge <- filter(fr_merge, year >= 2000 & year < 2019)
nl_merge <- filter(nl_merge, year >= 2000 & year < 2019)
pl_merge <- filter(pl_merge, year >= 2000 & year < 2019)

ggplot() +
  geom_line(data = ger_merge, aes(x = year, y = eu_empl_durtotal_nuts0, color = "black")) +
  geom_line(data = fr_merge, aes(x = year, y = eu_empl_durtotal_nuts0, color = "blue")) +
  geom_line(data = nl_merge, aes(x = year, y = eu_empl_durtotal_nuts0, color = "orange")) +
  geom_line(data = pl_merge, aes(x = year, y = eu_empl_durtotal_nuts0, color = "red")) +
  ggtitle("")
```

```{r rmployment rate }
library(tidyverse)

info <- read.csv("qog_eureg_wide1_nov20.csv")

gerinfo <- filter(info, region_name == "GERMANY")

spec <-  select(germaninfo, region_name, year)

ger_fil <- select(gerinfo, eu_emp_1524f_nuts0:eu_emp_2534t_nuts0)

ger_fil <- select(ger_fil, ends_with("nuts0"))

ger_merge <- bind_cols(spec, ger_fil)

ger_merge <- filter(ger_merge, year >= 1999 & year < 2019)

ggplot(data = ger_merge) +
  geom_line(mapping = aes(x = year, y = eu_emp_1524f_nuts0, color = "Female 15-24 years old")) +
  geom_line(mapping = aes(x = year, y = eu_emp_1524m_nuts0, color = "Male 15-24 years old")) +
  geom_line(mapping = aes(x = year, y = eu_emp_2534f_nuts0, color = "Female 25-34 years old")) +
  geom_line(mapping = aes(x = year, y = eu_emp_2534m_nuts0, color = "Male 25-34 years old")) +
  scale_color_manual(name = "Age Groups", values = c("Female 15-24 years old" = "#0000ff", "Male 15-24 years old" = "#7979ff","Female 25-34 years old" = "#e600ff", "Male 25-34 years old" = "#ff6ef8"))
```

```{r}
info <- read.csv("qog_eureg_wide1_nov20.csv")

gerinfo <- filter(info, region_name == "GERMANY")

```

```{r}
str(info)
```

```{r}
colnames(info)
```

 geom_line(mapping = aes(x = year, y = eu_emp_2064f_nuts0, color = "Female 20-60 years old")) +
  geom_line(mapping = aes(x = year, y = eu_emp_2064m_nuts0, color = "Male 20-60 years old")) +

   "Female 20-60 years old" = "#ff0000", "Male 20-60 years old" = "#ff7575", 