---
title: "Data visualisation"
author: "dr. Geert Monsieur"
output: html_document
---

# Data visualisation

## Prerequisites

One of the main libraries used in the textbook is tidyverse, which can be loaded by calling the `library` function.

```{r}
library(tidyverse)
```

## First steps

Do cars with big engines use more fuel than cars with small engines? You probably already have an answer.

### The `mpg` data frame

The data frame (i.e. the dataset) we use, is part of the tidyverse library and can be retrieved with `mpg`.

```{r}
mpg
```

Documentation can be consulted by adding a question mark to the data frame name, namely `?mpg`.

```{r}
?mpg
```

### Creating a ggplot

Plots we make in this course are typically created using the famous `ggplot` function. `ggplot` creates a coordinate system that you can add layers to. The first argument of the function is the dataset to use in the graph. Simply `ggplot(data = mpg)` creates an empty graph. Therefore, we add a first layer, which in this case is created by the `geom_point` function and renders a scatter plot.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Aesthetic mappings

The *color* of the points can be customized in the aesthetic mappings.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

The *size* of the points can be customized in the aesthetic mappings.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

The *transparancy* (i.e. alpha) of the points can be customized in the aesthetic mappings.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

The *shape* of the points can be customized in the aesthetic mappings.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

## Common problems

Pay attention to where you break the line when defining the layers in your `ggplot`-based plot.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```


## Facets

When you want to create multiple similar plots, facets allow to split your plot into _facets_, which are subplots that each display one subset of the data.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```

Similarly, you can also create grids of facets.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```


## Geometric objects

Trendlines can be added as an additional layer in a plot by calling the `geom_smooth` function.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

Please note ggplot is layered based and common settings can be set in the ggplot function.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes (color = class)) +
  geom_smooth()
```

## Statistical transformations

Bar plots can be created using the `geom_bar` function.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```
