# Tidy data

## Introduction

### Prerequisites

```{r setup, message = FALSE}
library(tidyverse)
```

## Tidy data

The example below shows the same data organised in four different ways. Each dataset shows the same values of four variables *country*, *year*, *population*, and *cases*, but each dataset organises the values in a different way.

```{r}
table1
table2
table3

# Spread across two tibbles
table4a  # cases
table4b  # population
```

Not really tidy, right? What do we need in orde have a _tidy_ dataset?

1.  Each variable must have its own column.
1.  Each observation must have its own row.
1.  Each value must have its own cell.

Hence, only `table1` is tidy. It's the only representation where each column is a variable.

Below a few examples showing how you might work with `table1`.

```{r, out.width = "75%"}
# Compute rate per 10,000
table1 %>% 
  mutate(rate = cases / population * 10000)

# Compute cases per year
table1 %>% 
  count(year, wt = cases)

# Visualise changes over time
library(ggplot2)
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```

## Pivoting

In many datasets you need to resolve one of two common problems:

1. One variable might be spread across multiple columns.

1. One observation might be scattered across multiple rows.

To fix these problems, you'll need the two most important functions in tidyr: `pivot_longer()` and `pivot_wider()`.

### Longer

A common problem is a dataset where some of the column names are not names of variables, but _values_ of a variable.

```{r}
table4a
```

To tidy a dataset like this, we need to __pivot__ the offending columns into a new pair of variables.

```{r}
table4a %>% 
  pivot_longer(c('1999', '2000'), names_to = "year", values_to = "cases")
```
Note that "1999" and "2000" are non-syntactic names (because they don't start with a letter) so we have to surround them in backticks.

`pivot_longer()` makes datasets _longer_ by _increasing the number of rows_ and decreasing the number of columns.

We can use `pivot_longer()` to tidy `table4b` in a similar fashion.

```{r}
table4b %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
```

To combine the tidied versions of `table4a` and `table4b` into a single tibble, we need to use `dplyr::left_join()`, which you'll learn about in Chapter 13.

```{r}
tidy4a <- table4a %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
tidy4b <- table4b %>% 
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
left_join(tidy4a, tidy4b)
```

### Wider

`pivot_wider()` is the opposite of `pivot_longer()`, which is what you need in `table2`, because each observation is spread across two rows.

```{r}
table2
```

Now we have to make the dataset _wider_:

```{r}
table2 %>%
    pivot_wider(names_from = type, values_from = count)
```

As you might have guessed from their names, `pivot_wider()` and `pivot_longer()` are complements. `pivot_longer()` makes wide tables narrower and longer; `pivot_wider()` makes long tables shorter and wider.

## Separating and uniting

### Separate

`table3` has a different problem: we have one column (`rate`) that contains two variables (`cases` and `population`). To fix this problem, we'll need the `separate()` function.

```{r}
table3
```

The `rate` column contains both `cases` and `population` variables, and we need to split it into two variables. `separate()` takes the name of the column to separate, and the names of the columns to separate into.

```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"))
```
By default, `separate()` will split values wherever it sees a non-alphanumeric character (i.e. a character that isn't a number or letter).

If you wish to use a specific character to separate a column, you can pass the character to the `sep` argument of `separate()`. Therefore, the following code does the same job.

```{r eval = FALSE}
table3 %>% 
  separate(rate, into = c("cases", "population"), sep = "/")
```

Did you notice the data type for both cases and population? The default behaviour in `separate()` is to leave the type of the column as is. Alternatively, you can ask `separate` to _convert_ to better types as follows:

```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"), convert = TRUE)
```

You can also specify the exact positions where to split at, using a vector of integers.


```{r}
table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)
```

### Unite

`unite()` is the inverse of `separate()`.

```{r}
table5 %>% 
  unite(new, century, year)
```

In this case we also need to use the `sep` argument. The default will place an underscore (`_`) between the values from different columns.

```{r}
table5 %>% 
  unite(new, century, year, sep = "")
```


## Missing values

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)

stocks
```

There are two missing values in this dataset!

* The return for the fourth quarter of 2015 is _explicitly_ missing, because
  the cell where its value should be instead contains `NA`.
  
* The return for the first quarter of 2016 is _implicitly_ missing, because it
  simply does not appear in the dataset.
  
`spread` helps us making the implicit missing value explicit by putting years in the columns:

```{r}
stocks %>% 
  spread(year, return)
```

Another important tool for making missing values explicit in tidy data is `complete()`:

```{r}
stocks %>% 
  complete(year, qtr)
```

`complete()` takes a set of columns, and finds all unique combinations. It then ensures the original dataset contains all those values, filling in explicit `NA`s where necessary.
